---
title: "Experimental sample analysis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Alexander McFarland"
output: 
  html_document:
    toc: true 
    toc_depth: 3  
    number_sections: true  
    theme: united  
    highlight: tango  
params:
  run_dir: ""
  experimental_trend_width: ""
  experimental_trend_height: ""
  per_experimental_group_width: ""
  per_experimental_group_height: ""
---

# Libraries and directories

This is a processing chunk.

```{r, load_libraries, echo=FALSE, message=FALSE, warning=FALSE}
options(scipen = 999)

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(tibble)
library(cowplot)

### Troubleshooting inputs start # hash out params in the yaml portion of Rmd to take effect
# run_dir <- data.frame(run_dir = '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-R21_tulane/bartender_output')
# run_dir <- data.frame(run_dir = '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-bnab_2_0/bartender_output')
# runs_paths <- data.frame(
#   'base' = run_dir,
#   'record' = file.path(run_dir, 'record', 'experimental.csv'),
#   'filtered_stock' = file.path(run_dir, 'stock', 'stock_barcodes_filtered.csv'),
#   'raw_stock' = file.path(run_dir, 'stock', 'stock_barcodes_raw.csv'),
#   'experimental' = file.path(run_dir, 'experimental'),
#   'stock' = file.path(run_dir, 'stock'),
#   'standard_barcode_names' = file.path(run_dir, 'output', 'standard_barcode_names.csv')
# )
# dynamic_figure_manager <- data.frame(
#   experimental_trend_width = as.numeric(15),
#   experimental_trend_height = as.numeric(15),
#   per_experimental_group_width = as.numeric(15),
#   per_experimental_group_height = as.numeric(15)
# )
### Troubleshooting inputs end

dynamic_figure_manager <- data.frame(
  experimental_trend_width = as.numeric(params$experimental_trend_width),
  experimental_trend_height = as.numeric(params$experimental_trend_height),
  per_experimental_group_width = as.numeric(params$per_experimental_group_width),
  per_experimental_group_height = as.numeric(params$per_experimental_group_height)
)

runs_paths <- data.frame(
  'base' = params$run_dir,
  'record' = file.path(params$run_dir, 'record', 'experimental.csv'),
  'filtered_stock' = file.path(params$run_dir, 'stock', 'stock_barcodes_filtered.csv'),
  'raw_stock' = file.path(params$run_dir, 'stock', 'stock_barcodes_raw.csv'),
  'experimental' = file.path(params$run_dir, 'experimental'),
  'stock' = file.path(params$run_dir, 'stock'),
  'standard_barcode_names' = file.path(params$run_dir, 'output', 'standard_barcode_names.csv')
)

df_record <- read.csv(file.path(runs_paths$record)) %>%
  dplyr::select(-c(R1, R2, sample_name, input_template)) %>%
  base::unique() %>%
  dplyr::mutate(cell_type = stringr::str_to_lower(cell_type))

keep_global_vars <- c(
  # Values
  'runs_paths',
  # Data
  'df_raw_stock',
  'df_filtered_stock',
  'df_record',
  'dynamic_figure_manager',
  # Self
  'keep_global_vars'
)

#' Remove time_point_description
get_unique_experimental_treatments <- function(df) {
  return(df %>%
    dplyr::select(-time_point_description) %>%
    base::unique())
  }
  
list_non_global_variables <- function(all_variables, global_variables_to_keep) {
  return(setdiff(all_variables, global_variables_to_keep))
}

internal_functions <- unlist(lapply(lsf.str(), function(x){x[1]}))

keep_global_vars <- c(keep_global_vars, internal_functions)

rm(list = list_non_global_variables(ls(), keep_global_vars))
```


## Barcodes trend

Samples are displayed in a heatmap, with the biological group on the Y axis and timepoint/cell type/organ/genetic material on the X-axis. If any of these are not available, they are listed as `not_specified`.

Each tile in the heatmap contains the `proportion of barcodes sourced to the stock`, followed by `unique barcodes found in stock` | `total barcodes found in stock`, followed by `unique barcodes not found in stock` | `total barcodes not found in stock`. Each tile is colored by the proportion of barcodes found in stock.

```{r, experimental_trend, message=FALSE, warning=FALSE, echo=FALSE, out.width = '100%',fig.align="center", fig.width = 15, fig.height=15}
rm(list = list_non_global_variables(ls(), keep_global_vars))

df_bg <- do.call(rbind, lapply(Sys.glob(file.path(runs_paths$experimental, '*_processed.csv')), function(file_) {
  df_sample <- read.csv(file_) %>%
    dplyr::mutate(standard_sample_name = stringr::str_replace(basename(file_), '_processed.csv', ''))}))

df_summary_bg <- df_bg %>%
  dplyr::filter(trusted_proportion > 0) %>%
  dplyr::mutate(in_stock = ifelse(instock_threshold_check == 'pass', TRUE, FALSE)) %>%
  dplyr::select(standard_sample_name, trusted_proportion, in_stock, total)

calculate_instock_trusted_proportion <- function(df) {
  df <- df %>%
    dplyr::select(standard_sample_name, trusted_proportion, in_stock) %>%
    dplyr::group_by(standard_sample_name, in_stock) %>%
    dplyr::mutate(trusted_proportion = sum(trusted_proportion)) %>%
    base::unique() %>%
    tidyr::pivot_wider(names_from = 'in_stock', values_from = 'trusted_proportion', values_fill = 0) %>%
    tidyr::pivot_longer(cols = -c(standard_sample_name), names_to = 'in_stock', values_to = 'trusted_proportion')
  return(df)
}

calculate_instock_total <- function(df) {
  df <- df %>%
    dplyr::select(standard_sample_name, total, in_stock) %>%
    dplyr::group_by(standard_sample_name, in_stock) %>%
    dplyr::mutate(total = sum(total)) %>%
    base::unique() %>%
    tidyr::pivot_wider(names_from = 'in_stock', values_from = 'total', values_fill = 0) %>%
    tidyr::pivot_longer(cols = -c(standard_sample_name), names_to = 'in_stock', values_to = 'total')
  return(df)
}

calculate_instock_unique <- function(df) {
  df <- df %>%
    dplyr::select(standard_sample_name, total, in_stock) %>%
    dplyr::group_by(standard_sample_name, in_stock) %>%
    dplyr::mutate(unique = length(total)) %>%
    dplyr::select(-total) %>%
    base::unique() %>%
    tidyr::pivot_wider(names_from = 'in_stock', values_from = 'unique', values_fill = 0) %>%
    tidyr::pivot_longer(cols = -c(standard_sample_name), names_to = 'in_stock', values_to = 'unique')
  return(df)
}

df_trusted_proportion <- calculate_instock_trusted_proportion(df_summary_bg)

df_total <- calculate_instock_total(df_summary_bg)

df_unique <- calculate_instock_unique(df_summary_bg)

df_plot <- merge(df_trusted_proportion, df_total, by = c('standard_sample_name', 'in_stock'))

df_plot <- merge(df_plot, df_unique, by = c('standard_sample_name', 'in_stock'))

df_plot <- merge(df_plot, df_record, by = 'standard_sample_name')

# make experimental treatment id
df_plot <- df_plot %>%
  dplyr::mutate(experimental_treatment_id = paste0(time_point, '_', cell_type, '_', organ, '_', genetic_source))

# keep only one experimental treatment per row
df_plot <- get_unique_experimental_treatments(df_plot)

df_plot_text <- df_plot %>%
  dplyr::mutate(tile_label = ifelse(in_stock == TRUE, paste(round(trusted_proportion, 2), '\n', scales::comma(unique), '|', scales::comma(total)), paste(scales::comma(unique), '|', scales::comma(total)))) %>%
  dplyr::group_by(standard_sample_name) %>%
  dplyr::arrange(dplyr::desc(in_stock)) %>%
  dplyr::mutate(tile_label2 = paste(tile_label, collapse = '\n')) %>%
  dplyr::ungroup() %>%
  dplyr::select(-c(in_stock, trusted_proportion, total, unique, tile_label)) %>%
  base::unique()


ggplot(df_plot %>%
         dplyr::filter(in_stock == TRUE),
       aes(x = experimental_treatment_id, y = biological_group, fill = trusted_proportion)) +
  ggh4x::facet_nested(biological_group ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free', switch = 'y') +
  geom_tile() +
  geom_text(data = df_plot_text %>% dplyr::mutate(trusted_proportion = 1), aes(label = tile_label2)) +
  theme_classic() +
  theme(
    strip.text.y.left = element_text(angle = 0),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = 'bottom'
    ) +
  scale_fill_gradient(low = 'orange', high = 'lightblue',
                      limits = c(0, 1), na.value = 'white') +
  labs(fill = 'proportion in stock')

```

## Barcodes detected by presence in stock

This section shows proportion of barcodes, the number of unique barcodes, and the total abundance of barcodes for each biological group according to whether they are detected in stock or not.

```{r, per_experimental_group, message=FALSE, warning=FALSE, echo=FALSE, out.width = '100%',fig.align="center", fig.width=dynamic_figure_manager$per_experimental_group_width, fig.height=dynamic_figure_manager$per_experimental_group_height}
rm(list = list_non_global_variables(ls(), keep_global_vars))

for (biological_group_ in base::unique(df_record$biological_group)) {

  df_bg <- do.call(rbind, lapply(Sys.glob(file.path(runs_paths$experimental, paste0(biological_group_, '*_processed.csv'))), function(file_) {
    df_sample <- read.csv(file_) %>%
      dplyr::mutate(standard_sample_name = stringr::str_replace(basename(file_), '_processed.csv', ''))

    return(df_sample)
  })
  )
  
  df_plot <- merge(df_bg, df_record, by = 'standard_sample_name')
  
  df_plot <- get_unique_experimental_treatments(df_plot)

  df_plot <- df_plot %>%
    dplyr::arrange(time_point) %>%
    dplyr::filter(trusted_proportion > 0) %>%
    dplyr::mutate(in_stock = ifelse(instock_threshold_check == 'pass', TRUE, FALSE))


  instock_fill_colors <- c('orange', 'lightblue')
  names(instock_fill_colors) <- c(FALSE, TRUE)

  plot_theme_ <- function(){
    theme_classic() +
    theme(
      strip.text = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.y = element_text(size = 15),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      plot.title = element_text(size = 15),
      legend.position = 'bottom'
    )
  }

  p1 <- ggplot(df_plot,
         aes(x = standard_sample_name, y = trusted_proportion, fill = in_stock)) +
    ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
    geom_bar(stat = 'identity', position = position_stack(), width = 1) +
    plot_theme_() +
    scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
    scale_x_discrete(expand = c(0, 0, 0, 0)) +
    labs(fill = 'in stock', y = 'proportion') +
    scale_fill_manual(values = instock_fill_colors) +
    ggtitle(label = paste(biological_group_, 'proportion barcodes'))

  p2 <- ggplot(df_plot,
         aes(x = standard_sample_name, y = total, fill = in_stock)) +
    ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
    geom_bar(stat = 'identity', position = position_stack(), width = 1) +
    plot_theme_() +
    scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
    scale_x_discrete(expand = c(0, 0, 0, 0)) +
    labs(fill = 'in stock', y = 'count') +
    scale_fill_manual(values = instock_fill_colors) +
    ggtitle(label = paste(biological_group_, 'total barcodes'))

  p3 <- ggplot(df_plot %>%
           dplyr::mutate(total = 1),
         aes(x = standard_sample_name, y = total, fill = in_stock)) +
    ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
    geom_bar(stat = 'identity', position = position_stack(), width = 1) +
    plot_theme_() +
    scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
    scale_x_discrete(expand = c(0, 0, 0, 0)) +
    labs(fill = 'in stock', y = 'count') +
    scale_fill_manual(values = instock_fill_colors) +
    ggtitle(label = paste(biological_group_, 'unique barcodes'))

  print(p1)

  print(p2)

  print(p3)

  }

```

## Barcodes detected by stock source

This section shows proportion of barcodes, the number of unique barcodes, and the total abundance of barcodes for each biological group according to a more refined view of the stock sourcing. Here, barcodes are color coded by what specific stock they were found in. If a barcode is not found in any stock, they are labeled as "experimental". If they are found in two or more different stocks then they are labeled "stock overlap". Otherwise, they are labeled with the name of stock.

```{r, per_experimental_group_specific_stock_source, message=FALSE, warning=FALSE, echo=FALSE, out.width = '100%',fig.align="center", fig.width=dynamic_figure_manager$per_experimental_group_width, fig.height=dynamic_figure_manager$per_experimental_group_height}
rm(list = list_non_global_variables(ls(), keep_global_vars))

df_barcode_source <- merge(
  read.csv(file.path(runs_paths$stock, 'stock_barcodes_filtered.csv')),
  read.csv(runs_paths$standard_barcode_names),
  by = 'barcode',
  all = TRUE)

df_barcode_source <- df_barcode_source %>%
  dplyr::select(barcode, standard_barcode_name, biological_group) %>%
  dplyr::mutate(stock_source = ifelse(stringr::str_detect(standard_barcode_name, 'stock'), 'stock', biological_group))

if (length(base::unique(read.csv(file.path(runs_paths$stock, 'stock_barcodes_filtered.csv'))$biological_group)) > 1) {
  
  for (biological_group_ in base::unique(df_record$biological_group)) {
  
    df_bg <- do.call(rbind, lapply(Sys.glob(file.path(runs_paths$experimental, paste0(biological_group_, '*_processed.csv'))), function(file_) {
      df_sample <- read.csv(file_) %>%
        dplyr::mutate(standard_sample_name = stringr::str_replace(basename(file_), '_processed.csv', ''))
  
      return(df_sample)
    })
    )
    
    df_plot <- merge(df_bg, df_record, by = 'standard_sample_name')
    
    df_plot <- merge(df_plot, df_barcode_source, by = 'barcode', all.x = TRUE)
    
    df_plot <- df_plot %>% 
      dplyr::mutate(
        stock_source = dplyr::case_when(
          stringr::str_detect(standard_barcode_name, 'stock') ~ 'stock_overlap',
          stringr::str_detect(standard_barcode_name, 'experimental') ~ 'experimental',
          TRUE ~ stock_source
        ))
  
    df_plot <- df_plot %>%
      dplyr::arrange(time_point) %>%
      dplyr::filter(trusted_proportion > 0)
    
    df_plot <- get_unique_experimental_treatments(df_plot)
  
    num_unique_stocks <- length(base::unique(df_plot$stock_source))
    instock_fill_colors <- c('orange', 'lightblue', RColorBrewer::brewer.pal(n = 8, name = 'Set3'))[1:num_unique_stocks]
    names(instock_fill_colors) <- c(c('experimental', 'stock_overlap'), base::unique(df_plot$stock_source)[!base::unique(df_plot$stock_source) %in% c('experimental', 'stock_overlap')])
    
    plot_theme_ <- function(){
      theme_classic() +
      theme(
        strip.text = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 15),
        legend.position = 'bottom'
      )
    }
  
    p1 <- ggplot(df_plot,
           aes(x = standard_sample_name, y = trusted_proportion, fill = stock_source)) +
      ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
      geom_bar(stat = 'identity', position = position_stack(), width = 1) +
      plot_theme_() +
      scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
      scale_x_discrete(expand = c(0, 0, 0, 0)) +
      labs(fill = 'source', y = 'proportion') +
      scale_fill_manual(values = instock_fill_colors) +
      ggtitle(label = paste(biological_group_, 'proportion barcodes'))
  
    p2 <- ggplot(df_plot,
           aes(x = standard_sample_name, y = total, fill = stock_source)) +
      ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
      geom_bar(stat = 'identity', position = position_stack(), width = 1) +
      plot_theme_() +
      scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
      scale_x_discrete(expand = c(0, 0, 0, 0)) +
      labs(fill = 'source', y = 'count') +
      scale_fill_manual(values = instock_fill_colors) +
      ggtitle(label = paste(biological_group_, 'total barcodes'))
  
    p3 <- ggplot(df_plot %>%
             dplyr::mutate(total = 1),
           aes(x = standard_sample_name, y = total, fill = stock_source)) +
      ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
      geom_bar(stat = 'identity', position = position_stack(), width = 1) +
      plot_theme_() +
      scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
      scale_x_discrete(expand = c(0, 0, 0, 0)) +
      labs(fill = 'source', y = 'count') +
      scale_fill_manual(values = instock_fill_colors) +
      ggtitle(label = paste(biological_group_, 'unique barcodes'))
  
    print(p1)
  
    print(p2)
  
    print(p3)
  
} else {
  print('Only one stock source used. This section will not be processed.')
}
  }

```