---
title: "Experimental sample analysis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Alexander McFarland"
output: 
  html_document:
    toc: true 
    toc_depth: 3  
    number_sections: true  
    theme: united  
    highlight: tango  
params:
  run_dir: ""
  experimental_trend_width: ""
  experimental_trend_height: ""
  per_experimental_group_width: ""
  per_experimental_group_height: ""
---

# Libraries and directories

This is a processing chunk.

```{r, load_libraries, echo=FALSE, message=FALSE, warning=FALSE}
options(scipen = 999)

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(tibble)
library(cowplot)

### Troubleshooting inputs start # hash out params in the yaml portion of Rmd to take effect
# run_dir <- data.frame(run_dir = '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-R21_tulane/bartender_output')
# run_dir <- data.frame(run_dir = '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-bnab_2_0/bartender_output')
# run_dir <- data.frame(run_dir = '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-bnab_at_arti/2_bnab_at_arti_animal_samples_stock')
# run_dir <- data.frame(run_dir = '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-bnab_at_arti/bartender_run_stock_and_experimental')
# run_dir <- data.frame(run_dir = '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-shiv_d_crispr/1_3-test_pipeline/animal_characterization_vs_shivd')
run_dir <- data.frame(run_dir = '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-shiv_d_crispr/1_4-test_pipeline/animal_characterization_vs_shivd')

runs_paths <- data.frame(
  'base' = run_dir,
  'record' = file.path(run_dir, 'record', 'experimental.csv'),
  'filtered_stock' = file.path(run_dir, 'stock', 'stock_barcodes_filtered.csv'),
  'raw_stock' = file.path(run_dir, 'stock', 'stock_barcodes_raw.csv'),
  'experimental' = file.path(run_dir, 'experimental'),
  'stock' = file.path(run_dir, 'stock'),
  'standard_barcode_names' = file.path(run_dir, 'output', 'standard_barcode_names.csv')
)
dynamic_figure_manager <- data.frame(
  experimental_trend_width = as.numeric(15),
  experimental_trend_height = as.numeric(15),
  per_experimental_group_width = as.numeric(15),
  per_experimental_group_height = as.numeric(15)
)
### Troubleshooting inputs end

# dynamic_figure_manager <- data.frame(
#   experimental_trend_width = as.numeric(params$experimental_trend_width),
#   experimental_trend_height = as.numeric(params$experimental_trend_height),
#   per_experimental_group_width = as.numeric(params$per_experimental_group_width),
#   per_experimental_group_height = as.numeric(params$per_experimental_group_height)
# )
# 
# runs_paths <- data.frame(
#   'base' = params$run_dir,
#   'record' = file.path(params$run_dir, 'record', 'experimental.csv'),
#   'filtered_stock' = file.path(params$run_dir, 'stock', 'stock_barcodes_filtered.csv'),
#   'raw_stock' = file.path(params$run_dir, 'stock', 'stock_barcodes_raw.csv'),
#   'experimental' = file.path(params$run_dir, 'experimental'),
#   'stock' = file.path(params$run_dir, 'stock'),
#   'standard_barcode_names' = file.path(params$run_dir, 'output', 'standard_barcode_names.csv')
# )

df_record <- read.csv(file.path(runs_paths$record)) %>%
  dplyr::select(-c(R1, R2, sample_name, input_template)) %>%
  base::unique() %>%
  dplyr::mutate(cell_type = stringr::str_to_lower(cell_type))

keep_global_vars <- c(
  # Values
  'runs_paths',
  # Data
  'df_raw_stock',
  'df_filtered_stock',
  'df_record',
  'dynamic_figure_manager',
  # Self
  'keep_global_vars'
)

#' Remove time_point_description
get_unique_experimental_treatments <- function(df) {
  return(df %>%
    dplyr::select(-time_point_description) %>%
    base::unique())
  }
  
list_non_global_variables <- function(all_variables, global_variables_to_keep) {
  return(setdiff(all_variables, global_variables_to_keep))
}

subchunkify <- function(g, fig_height=7, fig_width=5) {
  g_deparsed <- paste0(deparse(
    function() {g}
  ), collapse = '')

  sub_chunk <- paste0("
  `","``{r sub_chunk_", floor(runif(1) * 10000), ", fig.height=",
   fig_height, ", fig.width=", fig_width, ", echo=FALSE, warning=FALSE, message=FALSE}",
  "\n(", 
    g_deparsed
    , ")()",
  "\n`","``
  ")

  cat(knitr::knit(text = knitr::knit_expand(text = sub_chunk), quiet = TRUE))
}

format_thousands <- function(x, digits = 1) {
  sapply(x, function(num) {
    if (abs(num) >= 1000) {
      paste0(round(num / 1000, digits), "k")
    } else {
      as.character(num)
    }
  })
}

markdown_header <- function(hashes = '##', text) {
  cat("\n\n", hashes, " ", text, "\n\n", sep = "")
}




internal_functions <- unlist(lapply(lsf.str(), function(x){x[1]}))

keep_global_vars <- c(keep_global_vars, internal_functions)

rm(list = list_non_global_variables(ls(), keep_global_vars))
```


# Barcodes trend

Samples are displayed in a heatmap, with the biological group on the Y axis and timepoint/cell type/organ/genetic material on the X-axis. If any of these are not available, they are listed as `not_specified`.

Each tile in the heatmap has the following layout:

| Percentage of barcodes sourced to the stock |                      |
|---------------------------------------------|----------------------|
| Unique barcodes found in stock              \|| Total barcodes found in stock  |
| Unique barcodes not found in stock          \|| Total barcodes not found in stock |

For clarity of reading, all numbers are rounded to the nearest whole number. Numbers greater than a thousand have a K appended to them.


<!-- `percentage of barcodes sourced to the stock` -->

<!-- `unique barcodes found in stock` | `total barcodes found in stock` -->

<!-- `unique barcodes not found in stock` | `total barcodes not found in stock` -->

Each tile is colored by the proportion of barcodes found in stock.

For mixed stocks, this only provides the statistic for a barcode being in **any** stock. In the plots below there are additional deconstructions of stock sourcing.

```{r, experimental_trend, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
rm(list = list_non_global_variables(ls(), keep_global_vars))

df_bg <- do.call(rbind, lapply(Sys.glob(file.path(runs_paths$experimental, '*_processed.csv')), function(file_) {
  df_sample <- read.csv(file_) %>%
    dplyr::mutate(standard_sample_name = stringr::str_replace(basename(file_), '_processed.csv', ''))}))

df_summary_bg <- df_bg %>%
  dplyr::filter(trusted_proportion > 0) %>%
  dplyr::mutate(in_stock = ifelse(instock_threshold_check == 'pass', TRUE, FALSE)) %>%
  dplyr::select(standard_sample_name, trusted_proportion, in_stock, total)

calculate_instock_trusted_proportion <- function(df) {
  df <- df %>%
    dplyr::select(standard_sample_name, trusted_proportion, in_stock) %>%
    dplyr::group_by(standard_sample_name, in_stock) %>%
    dplyr::mutate(trusted_proportion = sum(trusted_proportion)) %>%
    base::unique() %>%
    tidyr::pivot_wider(names_from = 'in_stock', values_from = 'trusted_proportion', values_fill = 0) %>%
    tidyr::pivot_longer(cols = -c(standard_sample_name), names_to = 'in_stock', values_to = 'trusted_proportion')
  return(df)
}

calculate_instock_total <- function(df) {
  df <- df %>%
    dplyr::select(standard_sample_name, total, in_stock) %>%
    dplyr::group_by(standard_sample_name, in_stock) %>%
    dplyr::mutate(total = sum(total)) %>%
    base::unique() %>%
    tidyr::pivot_wider(names_from = 'in_stock', values_from = 'total', values_fill = 0) %>%
    tidyr::pivot_longer(cols = -c(standard_sample_name), names_to = 'in_stock', values_to = 'total')
  return(df)
}

calculate_instock_unique <- function(df) {
  df <- df %>%
    dplyr::select(standard_sample_name, total, in_stock) %>%
    dplyr::group_by(standard_sample_name, in_stock) %>%
    dplyr::mutate(unique = length(total)) %>%
    dplyr::select(-total) %>%
    base::unique() %>%
    tidyr::pivot_wider(names_from = 'in_stock', values_from = 'unique', values_fill = 0) %>%
    tidyr::pivot_longer(cols = -c(standard_sample_name), names_to = 'in_stock', values_to = 'unique')
  return(df)
}

df_trusted_proportion <- calculate_instock_trusted_proportion(df_summary_bg)

df_total <- calculate_instock_total(df_summary_bg)

df_unique <- calculate_instock_unique(df_summary_bg)

df_plot <- merge(df_trusted_proportion, df_total, by = c('standard_sample_name', 'in_stock'))

df_plot <- merge(df_plot, df_unique, by = c('standard_sample_name', 'in_stock'))

df_plot <- merge(df_plot, df_record, by = 'standard_sample_name')

# make experimental treatment id
df_plot <- df_plot %>%
  dplyr::mutate(experimental_treatment_id = paste0(time_point, '_', cell_type, '_', organ, '_', genetic_source))

# keep only one experimental treatment per row
df_plot <- get_unique_experimental_treatments(df_plot)

df_plot_text <- df_plot %>%
  # dplyr::mutate(
  #   tile_label = ifelse(in_stock == TRUE, paste(round(trusted_proportion, 2), '\n', scales::comma(unique), '|', scales::comma(total)), paste(scales::comma(unique), '|', scales::comma(total)))
  #   ) %>%
  # 
  dplyr::mutate(
    tile_label = ifelse(in_stock == TRUE, paste0(paste0(100 * round(trusted_proportion, 2),'%'), '\n', format_thousands(unique, 0), '|', format_thousands(total, 0)), paste0(format_thousands(unique, 0), '|', format_thousands(total, 0)))
    ) %>%
  dplyr::group_by(standard_sample_name) %>%
  dplyr::arrange(dplyr::desc(in_stock)) %>%
  dplyr::mutate(tile_label2 = paste(tile_label, collapse = '\n')) %>%
  dplyr::ungroup() %>%
  dplyr::select(-c(in_stock, trusted_proportion, total, unique, tile_label)) %>%
  base::unique()


p1 <- ggplot(df_plot %>%
         dplyr::filter(in_stock == TRUE),
       aes(x = experimental_treatment_id, y = biological_group, fill = trusted_proportion)) +
  ggh4x::facet_nested(biological_group ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free', switch = 'y') +
  geom_tile() +
  geom_text(data = df_plot_text %>% dplyr::mutate(trusted_proportion = 1), aes(label = tile_label2)) +
  theme_classic() +
  theme(
    strip.text.y.left = element_text(angle = 0),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = 'bottom'
    ) +
  scale_fill_gradient(low = 'orange', high = 'lightblue',
                      limits = c(0, 1), na.value = 'white') +
  labs(fill = 'proportion in stock')

fig_width_ <- p1$data %>% dplyr::select(time_point, cell_type, organ, genetic_source) %>%
  base::unique() %>% nrow()

fig_width_ <- (fig_width_/1.5) + 10 #sapply(seq(1, 50), function(x) {(x /1.5) + 10})

fig_height_ <- p1$data %>% dplyr::select(biological_group) %>%
  base::unique() %>% nrow()

fig_height_ <- (fig_height_/1.5)+5 #sapply(seq(1, 50), function(x) {(x /1.5) + 5})

subchunkify(p1, fig_width = fig_width_, fig_height = fig_height_)

```

# Barcode size distribution

```{r}

df_bg <- do.call(dplyr::bind_rows, lapply(Sys.glob(file.path(runs_paths$experimental, '*_barcode.csv')), function(file_) {
  df_sample <- read.csv(file_)
    # print(file.exists(file_))
    # dplyr::mutate(standard_sample_name = stringr::str_replace(basename(file_), '_barcode.csv', ''))
  }
  )
  ) %>%
  tidyr::pivot_longer(names_to = 'standard_sample_name', values_to = 'count', cols = -barcode) %>%
  dplyr::mutate(
    standard_sample_name = gsub("_[^_]*$", "", standard_sample_name),
    barcode_length = nchar(barcode)
    ) %>%
  dplyr::filter(!is.na(count)) %>%
  dplyr::group_by(barcode) %>%
  dplyr::mutate(samples_obs = dplyr::n_distinct(standard_sample_name)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(standard_sample_name) %>%
  dplyr::mutate(
    total_bc = sum(count)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(standard_sample_name, barcode) %>%
  dplyr::mutate(barcode_percentage = 100 * (count/total_bc)) %>%
  dplyr::ungroup()

df_barcode_source <- read.csv(file.path(runs_paths$stock, 'stock_barcodes_filtered.csv'))

df_barcode_source_raw <- read.csv(file.path(runs_paths$stock, 'stock_barcodes_raw.csv'))

df_v2_stock <- read.csv(
  file.path(
    '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-bnab_at_arti/4_1-cd4_source_troubleshooting/bartender_mixed_stock/stock',
    'stock_barcodes_raw.csv'
    )
  )

df_subset <- df_bg %>%
  dplyr::filter(barcode_length > 30 & barcode_length < 40) %>%
  dplyr::left_join(
    df_record, 
    by = 'standard_sample_name'
  ) %>%
  dplyr::left_join(
    df_barcode_source %>%
      dplyr::select(biological_group, barcode) %>%
      dplyr::rename(stock_id = biological_group),
    by = 'barcode', 
  ) %>%
  dplyr::mutate(
    in_barcode_raw = barcode %in% df_barcode_source_raw$barcode,
    in_v2_barcode_raw = barcode %in% df_v2_stock$barcode
  )

df_plot <- df_subset %>%
  dplyr::group_by(standard_sample_name, barcode_length) %>%
  dplyr::mutate(total_percentage_at_length = sum(barcode_percentage)) %>%
  dplyr::ungroup() %>%
  dplyr::select(standard_sample_name, barcode_length, total_percentage_at_length) %>%
  dplyr::distinct() %>%
  dplyr::left_join(
    df_record,
    by = 'standard_sample_name'
  )

colors <- c(
  "#EE6677", # 30
  "#66CCEE",
  "#228833",
  "#CCBB44",
  "#EE6677",
  "#4477AA",
  "#BBBBBB",
  "#000000",
  "#44AA99",
  "#EE8866"
)

bc_fill <- colors
names(bc_fill) <- as.character(seq(30, 39, by = 1))

# whats the summed proportion of barcodes by length
ggplot(df_plot, aes(fill = as.factor(barcode_length), y = total_percentage_at_length, x = standard_sample_name)) +
  ggh4x::facet_nested(. ~ cell_type + biological_group + organ, scales = 'free', space = 'free', switch = 'y') +
  geom_bar(stat = 'identity', position = position_stack()) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 5)
  ) +
  scale_fill_manual(values = bc_fill) +
  labs(fill = 'bc len')


df_bc_dom <- tibble::tribble(
  ~standard_sample_name, ~barcode_dominance,
  "DO52_CD4_0_SM_DNA", "34_dom",
  "DP55_CD4_0_SM_DNA", "34_dom",
  "DO58_CD4_0_SM_DNA", "34_dom",
  "DP12_CD4_0_SM_DNA", "34_dom",
  "DR30_CD4_0_SM_DNA", "34_dom",
  "DM58_CD4_0_SM_DNA", "34_dom",
  "DP66_MACROPHAGE_0_SM_DNA", "34_dom",
  
  
  "DO60_MACROPHAGE_0_SM_DNA", "35_dom",
  "MD86_CD4_0_NB_DNA", "35_dom",
  "KH22_CD4_0_NB_DNA", "35_dom",
  "ML02_CD4_0_SM_DNA", "35_dom",
  "DO52_MACROPHAGE_0_SM_DNA", "35_dom",
  "ML02_CD4_0_NB_DNA", "35_dom",
  
  "Not_Specified_CD4_0_not_applicable_DNA", "other"
  )




  
for (i_ in c(34, 35)) {  
  # i_ <- 34
  
  # sum of the five largest barcodes of size 34 and 35  
  df_plot <- df_subset %>%
    dplyr::filter(barcode_length == i_) %>%
    dplyr::group_by(standard_sample_name, barcode_length) %>%
    dplyr::slice_max(order_by = barcode_percentage, n = 5) %>%
    dplyr::arrange(dplyr::desc(barcode_percentage)) %>%
    dplyr::mutate(barcode_order = seq(1, dplyr::n())) %>%
    dplyr::ungroup() %>%
    dplyr::left_join(
      df_bc_dom,
      by = 'standard_sample_name'
    ) %>%
    dplyr::mutate(standard_sample_name = factor(standard_sample_name, levels = df_bc_dom$standard_sample_name))
  
  p1 <- ggplot(df_plot, aes(color = standard_sample_name, y = barcode_percentage, x = barcode_order)) +
    ggh4x::facet_nested(barcode_dominance ~ ., scales = 'free', space = 'fixed') +
    geom_line() +
    geom_point() +
    theme_light() +
    theme(
      axis.text.x = element_text(size = 10)
    ) +
    labs(title = paste0('top 5 BC of length ', i_))
  
  print(p1)
  
  # total numbers of barcodes
  df_plot <- df_subset %>%
    dplyr::filter(barcode_length == i_) %>%
    dplyr::group_by(standard_sample_name) %>%
    dplyr::mutate(total_bc_of_len = dplyr::n_distinct(barcode)) %>%
    dplyr::ungroup() %>%
    dplyr::select(standard_sample_name, total_bc_of_len) %>%
    dplyr::distinct() %>%
    dplyr::left_join(
      df_bc_dom,
      by = 'standard_sample_name'
    ) %>%
    dplyr::mutate(standard_sample_name = factor(standard_sample_name, levels = df_bc_dom$standard_sample_name))
  
  p2 <- ggplot(df_plot, aes(y = standard_sample_name, x = total_bc_of_len, fill = standard_sample_name)) +
    ggh4x::facet_nested(barcode_dominance ~ ., scales = 'free', space = 'fixed') +
    geom_bar(stat = 'identity') +
    theme_light() +
    theme(
      axis.text.x = element_text(size = 10),
      axis.text.y = element_text(size = 10),
      legend.position = 'none',
      axis.title.y = element_blank()
    ) +
    labs(title = paste0('Number unique barcodes of length ', i_))
  
  print(p2)
  
  # total numbers of barcodes split by found in SHIV D stock raw
  df_plot <- df_subset %>%
    dplyr::filter(barcode_length == i_) %>%
    dplyr::mutate(
      barcode_source = dplyr::case_when(
        (in_barcode_raw == T & in_v2_barcode_raw == T) ~ 'both',
        (in_barcode_raw == T & in_v2_barcode_raw == F) ~ 'SHIVD',
        (in_barcode_raw == F & in_v2_barcode_raw == T) ~ 'V2',
        (in_barcode_raw == F & in_v2_barcode_raw == F) ~ 'neither',
        TRUE ~ 'lookatme'
      )
    ) %>%
    dplyr::group_by(standard_sample_name, barcode_source) %>%
    dplyr::mutate(total_bc_of_len = dplyr::n_distinct(barcode)) %>%
    dplyr::ungroup() %>%
    dplyr::select(standard_sample_name, total_bc_of_len, barcode_source) %>%
    dplyr::distinct() %>%
    dplyr::left_join(
      df_bc_dom,
      by = 'standard_sample_name'
    ) %>%
    dplyr::mutate(
      standard_sample_name = factor(standard_sample_name, levels = df_bc_dom$standard_sample_name),
      barcode_source = factor(
        barcode_source,
        levels = c('both', 'SHIVD', 'V2', 'neither')
      )
      )
  
  p3 <- ggplot(df_plot, aes(y = standard_sample_name, x = total_bc_of_len, fill = barcode_source)) +
    ggh4x::facet_nested(barcode_dominance ~ ., scales = 'free', space = 'fixed') +
    geom_bar(stat = 'identity') +
    theme_light() +
    theme(
      axis.text.x = element_text(size = 10),
      axis.text.y = element_text(size = 10),
      legend.position = 'bottom',
      axis.title.y = element_blank()
    ) +
    labs(title = paste0('Number unique barcodes of length ', i_)) +
    scale_fill_brewer(palette = 'Set2') +
    guides(fill = guide_legend(reverse = T))
  
  print(p3)
   
  }



    

```


# Shivd cns Barcode size distribution

```{r}

shivd_cns_path <- '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-SHIV_D_CNS/bartender_run_stock_and_experimental_3'

shivd_cns_record <- read.csv(file.path(shivd_cns_path, 'record', 'experimental.csv')) %>%
  dplyr::filter((cell_type == 'CD4_T_CELL' & organ == 'PBMC') | (cell_type == 'MACROPHAGE' & organ == 'PBMC') |
                  (cell_type == 'PLASMA' & time_point == 7 & organ == 'BLOOD')) %>%
  dplyr::select(-c(sample_name, R1, R2, input_template)) %>%
  dplyr::distinct()

df_barcode_source <- read.csv(file.path(runs_paths$stock, 'stock_barcodes_filtered.csv'))

df_barcode_source_raw <- read.csv(file.path(runs_paths$stock, 'stock_barcodes_raw.csv'))

df_v2_stock <- read.csv(
  file.path(
    '/data/bar_lab/small_scale_barcode_analyses/data/interim/1-bnab_at_arti/4_1-cd4_source_troubleshooting/bartender_mixed_stock/stock',
    'stock_barcodes_raw.csv'
    )
  )

df_bg <- do.call(dplyr::bind_rows, lapply(Sys.glob(file.path(shivd_cns_path, 'experimental', paste0(shivd_cns_record$standard_sample_name, '_barcode.csv'))), function(file_) {
  df_sample <- read.csv(file_)
    # print(file.exists(file_))
    # dplyr::mutate(standard_sample_name = stringr::str_replace(basename(file_), '_barcode.csv', ''))
  }
  )
  ) %>%
  tidyr::pivot_longer(names_to = 'standard_sample_name', values_to = 'count', cols = -barcode) %>%
  dplyr::mutate(
    standard_sample_name = gsub("_[^_]*$", "", standard_sample_name),
    barcode_length = nchar(barcode)
    ) %>%
  dplyr::filter(!is.na(count)) %>%
  dplyr::group_by(barcode) %>%
  dplyr::mutate(samples_obs = dplyr::n_distinct(standard_sample_name)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(standard_sample_name) %>%
  dplyr::mutate(
    total_bc = sum(count)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(standard_sample_name, barcode) %>%
  dplyr::mutate(barcode_percentage = 100 * (count/total_bc)) %>%
  dplyr::ungroup()



colors <- c(
  "#EE6677", # 30
  "#66CCEE",
  "#228833",
  "#CCBB44",
  "#EE6677",
  "#4477AA",
  "#BBBBBB",
  "#000000",
  "#44AA99",
  "#EE8866"
)

bc_fill <- colors
names(bc_fill) <- as.character(seq(30, 39, by = 1))

for (cell_type_ in c('PLASMA', 'MACROPHAGE', 'CD4_T_CELL')) {
  
  df_subset <- df_bg %>%
    dplyr::filter(barcode_length > 30 & barcode_length < 40) %>%
    dplyr::left_join(
      shivd_cns_record, 
      by = 'standard_sample_name'
    ) %>%
    dplyr::filter(cell_type == cell_type_) %>%
    dplyr::mutate(
      in_barcode_raw = barcode %in% df_barcode_source_raw$barcode,
      in_v2_barcode_raw = barcode %in% df_v2_stock$barcode
    )
  
  df_plot <- df_subset %>%
    dplyr::group_by(standard_sample_name, barcode_length) %>%
    dplyr::mutate(total_percentage_at_length = sum(barcode_percentage)) %>%
    dplyr::ungroup() %>%
    dplyr::select(standard_sample_name, barcode_length, total_percentage_at_length) %>%
    dplyr::distinct() %>%
    dplyr::left_join(
      shivd_cns_record,
      by = 'standard_sample_name'
    )

  # whats the summed proportion of barcodes by length
  p1 <- ggplot(df_plot, aes(fill = as.factor(barcode_length), y = total_percentage_at_length, x = standard_sample_name)) +
    ggh4x::facet_nested(. ~ cell_type + biological_group + organ, scales = 'free', space = 'free', switch = 'y') +
    geom_bar(stat = 'identity', position = position_stack()) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 5)
    ) +
    scale_fill_manual(values = bc_fill) +
    labs(fill = 'bc len', title = paste0('Cell type: ', cell_type_))
  
  print(p1)


  
  # TBD
  df_bc_dom <- df_plot %>%
    dplyr::filter(barcode_length %in% c(34, 35)) %>%
    dplyr::group_by(standard_sample_name) %>%
    dplyr::slice_max(order_by = total_percentage_at_length, with_ties = F) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      barcode_dominance = ifelse(barcode_length == 34, '34_dom', '35_dom')
    )
  
  
  for (i_ in c(34, 35)) {  
    # i_ <- 34
    
    # sum of the five largest barcodes of size 34 and 35  
    df_plot <- df_subset %>%
      dplyr::filter(barcode_length == i_) %>%
      dplyr::group_by(standard_sample_name, barcode_length) %>%
      dplyr::slice_max(order_by = barcode_percentage, n = 5) %>%
      dplyr::arrange(dplyr::desc(barcode_percentage)) %>%
      dplyr::mutate(barcode_order = seq(1, dplyr::n())) %>%
      dplyr::ungroup() %>%
      dplyr::left_join(
        df_bc_dom,
        by = 'standard_sample_name'
      ) %>%
      dplyr::mutate(standard_sample_name = factor(standard_sample_name, levels = df_bc_dom$standard_sample_name))
    
    p1 <- ggplot(df_plot, aes(color = standard_sample_name, y = barcode_percentage, x = barcode_order)) +
      ggh4x::facet_nested(barcode_dominance ~ ., scales = 'free', space = 'fixed') +
      geom_line() +
      geom_point() +
      theme_light() +
      theme(
        axis.text.x = element_text(size = 10)
      ) +
      labs(title = paste0('Cell type: ', cell_type_, '\nTop 5 BC of length ', i_))
    
    print(p1)
    
    # total numbers of barcodes split by found in SHIV D stock raw
    df_plot <- df_subset %>%
      dplyr::filter(barcode_length == i_) %>%
      dplyr::mutate(
        barcode_source = dplyr::case_when(
          (in_barcode_raw == T & in_v2_barcode_raw == T) ~ 'both',
          (in_barcode_raw == T & in_v2_barcode_raw == F) ~ 'SHIVD',
          (in_barcode_raw == F & in_v2_barcode_raw == T) ~ 'V2',
          (in_barcode_raw == F & in_v2_barcode_raw == F) ~ 'neither',
          TRUE ~ 'lookatme'
        )
      ) %>%
      dplyr::group_by(standard_sample_name, barcode_source) %>%
      dplyr::mutate(total_bc_of_len = dplyr::n_distinct(barcode)) %>%
      dplyr::ungroup() %>%
      dplyr::select(standard_sample_name, total_bc_of_len, barcode_source) %>%
      dplyr::distinct() %>%
      dplyr::left_join(
        df_bc_dom,
        by = 'standard_sample_name'
      ) %>%
      dplyr::mutate(
        standard_sample_name = factor(standard_sample_name, levels = df_bc_dom$standard_sample_name),
        barcode_source = factor(
          barcode_source,
          levels = c('both', 'SHIVD', 'V2', 'neither')
        )
        )
    
    p3 <- ggplot(df_plot, aes(y = standard_sample_name, x = total_bc_of_len, fill = barcode_source)) +
      ggh4x::facet_nested(barcode_dominance ~ ., scales = 'free', space = 'fixed') +
      geom_bar(stat = 'identity') +
      theme_light() +
      theme(
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = 'bottom',
        axis.title.y = element_blank()
      ) +
      labs(title = paste0('Cell type: ', cell_type_, '\nNumber unique barcodes of length ', i_)) +
      scale_fill_brewer(palette = 'Set2') +
      guides(fill = guide_legend(reverse = T))
    
    print(p3)
     
  }
  
  }



    

```

<!-- # Barcodes detected by presence in stock -->

<!-- This section shows proportion of barcodes, the number of unique barcodes, and the total abundance of barcodes for each biological group according to whether they are detected in stock or not. -->

<!-- ```{r, per_experimental_group, message=FALSE, warning=FALSE, echo=FALSE, out.width = '100%',fig.align="center", fig.width=dynamic_figure_manager$per_experimental_group_width, fig.height=dynamic_figure_manager$per_experimental_group_height} -->
<!-- rm(list = list_non_global_variables(ls(), keep_global_vars)) -->

<!-- for (biological_group_ in base::unique(df_record$biological_group)) { -->

<!--   df_bg <- do.call(rbind, lapply(Sys.glob(file.path(runs_paths$experimental, paste0(biological_group_, '*_processed.csv'))), function(file_) { -->
<!--     df_sample <- read.csv(file_) %>% -->
<!--       dplyr::mutate(standard_sample_name = stringr::str_replace(basename(file_), '_processed.csv', '')) -->

<!--     return(df_sample) -->
<!--   }) -->
<!--   ) -->

<!--   df_plot <- merge(df_bg, df_record, by = 'standard_sample_name') -->

<!--   df_plot <- get_unique_experimental_treatments(df_plot) -->

<!--   df_plot <- df_plot %>% -->
<!--     dplyr::arrange(time_point) %>% -->
<!--     dplyr::filter(trusted_proportion > 0) %>% -->
<!--     dplyr::mutate(in_stock = ifelse(instock_threshold_check == 'pass', TRUE, FALSE)) -->


<!--   instock_fill_colors <- c('orange', 'lightblue') -->
<!--   names(instock_fill_colors) <- c(FALSE, TRUE) -->

<!--   plot_theme_ <- function(){ -->
<!--     theme_classic() + -->
<!--     theme( -->
<!--       strip.text = element_text(size = 15), -->
<!--       axis.text.y = element_text(size = 15), -->
<!--       axis.title.y = element_text(size = 15), -->
<!--       axis.text.x = element_blank(), -->
<!--       axis.title.x = element_blank(), -->
<!--       plot.title = element_text(size = 15), -->
<!--       legend.position = 'bottom' -->
<!--     ) -->
<!--   } -->

<!--   p1 <- ggplot(df_plot, -->
<!--          aes(x = standard_sample_name, y = trusted_proportion, fill = in_stock)) + -->
<!--     ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') + -->
<!--     geom_bar(stat = 'identity', position = position_stack(), width = 1) + -->
<!--     plot_theme_() + -->
<!--     scale_y_continuous(expand = c(0, 0, 0.1, 0)) + -->
<!--     scale_x_discrete(expand = c(0, 0, 0, 0)) + -->
<!--     labs(fill = 'in stock', y = 'proportion') + -->
<!--     scale_fill_manual(values = instock_fill_colors) + -->
<!--     ggtitle(label = paste(biological_group_, 'proportion barcodes')) -->

<!--   p2 <- ggplot(df_plot, -->
<!--          aes(x = standard_sample_name, y = total, fill = in_stock)) + -->
<!--     ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') + -->
<!--     geom_bar(stat = 'identity', position = position_stack(), width = 1) + -->
<!--     plot_theme_() + -->
<!--     scale_y_continuous(expand = c(0, 0, 0.1, 0)) + -->
<!--     scale_x_discrete(expand = c(0, 0, 0, 0)) + -->
<!--     labs(fill = 'in stock', y = 'count') + -->
<!--     scale_fill_manual(values = instock_fill_colors) + -->
<!--     ggtitle(label = paste(biological_group_, 'total barcodes')) -->

<!--   p3 <- ggplot(df_plot %>% -->
<!--            dplyr::mutate(total = 1), -->
<!--          aes(x = standard_sample_name, y = total, fill = in_stock)) + -->
<!--     ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') + -->
<!--     geom_bar(stat = 'identity', position = position_stack(), width = 1) + -->
<!--     plot_theme_() + -->
<!--     scale_y_continuous(expand = c(0, 0, 0.1, 0)) + -->
<!--     scale_x_discrete(expand = c(0, 0, 0, 0)) + -->
<!--     labs(fill = 'in stock', y = 'count') + -->
<!--     scale_fill_manual(values = instock_fill_colors) + -->
<!--     ggtitle(label = paste(biological_group_, 'unique barcodes')) -->

<!--   print(p1) -->

<!--   print(p2) -->

<!--   print(p3) -->

<!--   } -->

<!-- ``` -->

# Barcodes detected by stock source

This section shows proportion of barcodes, the number of unique barcodes, and the total abundance of barcodes for each biological group according to a more refined view of the stock sourcing. Here, barcodes are color coded by what specific stock they were found in. If a barcode is not found in any stock, they are labeled as "experimental". If they are found in two or more different stocks then they are labeled "stock overlap". Otherwise, they are labeled with the name of stock.

<!--# ```{r, per_experimental_group_specific_stock_source, message=FALSE, warning=FALSE, echo=FALSE,fig.align="center", fig.width=6, fig.height=6}-->
```{r, per_experimental_group_specific_stock_source, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
rm(list = list_non_global_variables(ls(), keep_global_vars))

plot_text_size_ <- 10

df_barcode_source <- merge(
  read.csv(file.path(runs_paths$stock, 'stock_barcodes_filtered.csv')),
  read.csv(runs_paths$standard_barcode_names),
  by = 'barcode',
  all = TRUE)

df_barcode_source <- df_barcode_source %>%
  dplyr::select(barcode, standard_barcode_name, biological_group) %>%
  dplyr::mutate(stock_source = ifelse(stringr::str_detect(standard_barcode_name, 'stock'), 'stock', biological_group))


unique_stock_names <- base::unique(df_barcode_source$stock_source)
unique_stock_names <- unique_stock_names[!is.na(unique_stock_names)]
unique_stock_names <- unique_stock_names[!unique_stock_names=='stock']
unique_stock_names <- base::unique(c('stock_overlap', 'experimental', unique_stock_names))
num_unique_stocks <- length(unique_stock_names)
instock_fill_colors <- c('orange', 'lightblue', RColorBrewer::brewer.pal(n = 8, name = 'Set3'))[1:num_unique_stocks]
names(instock_fill_colors) <- unique_stock_names


barcode_stock_source <- list()


# if (10 > 1) {
if (length(base::unique(read.csv(file.path(runs_paths$stock, 'stock_barcodes_filtered.csv'))$biological_group)) > 1) {

  for (biological_group_ in base::unique(df_record$biological_group)) {

    df_bg <- do.call(rbind, lapply(Sys.glob(file.path(runs_paths$experimental, paste0(biological_group_, '*_processed.csv'))), function(file_) {
      df_sample <- read.csv(file_) %>%
        dplyr::mutate(standard_sample_name = stringr::str_replace(basename(file_), '_processed.csv', ''))

      return(df_sample)
    })
    )
    
    if (is.null(df_bg)) {
      next
    }

    df_plot <- merge(df_bg, df_record, by = 'standard_sample_name')

    df_plot <- merge(df_plot, df_barcode_source, by = 'barcode', all.x = TRUE)

    df_plot <- df_plot %>%
      dplyr::mutate(
        stock_source = dplyr::case_when(
          stringr::str_detect(standard_barcode_name, 'stock') ~ 'stock_overlap',
          stringr::str_detect(standard_barcode_name, 'experimental') ~ 'experimental',
          TRUE ~ stock_source
        ))

    df_plot <- df_plot %>%
      dplyr::arrange(time_point) %>%
      dplyr::filter(trusted_proportion > 0)

    df_plot <- get_unique_experimental_treatments(df_plot)

    # unique_stock_names <- base::unique(df_plot$stock_source)
    # unique_stock_names <- base::unique(c('stock_overlap', 'experimental', unique_stock_names))
    # num_unique_stocks <- length(unique_stock_names)
    # instock_fill_colors <- c('orange', 'lightblue', RColorBrewer::brewer.pal(n = 8, name = 'Set3'))[1:num_unique_stocks]
    # names(instock_fill_colors) <- unique_stock_names

    plot_theme_ <- function(){
      theme_classic() +
      theme(
        text = element_text(size = plot_text_size_),
        # strip.text = element_text(size = plot_text_size_),
        # axis.text.y = element_text(size = plot_text_size_),
        # axis.title.y = element_text(size = plot_text_size_),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        # plot.title = element_text(size = plot_text_size_),
        legend.position = 'bottom',
        legend.title=element_blank()
      )
    }

    p1 <- ggplot(df_plot,
           aes(x = standard_sample_name, y = trusted_proportion, fill = stock_source)) +
      ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
      geom_bar(stat = 'identity', position = position_stack(), width = 1) +
      plot_theme_() +
      scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
      scale_x_discrete(expand = c(0, 0, 0, 0)) +
      labs(fill = 'source', y = 'proportion') +
      scale_fill_manual(values = instock_fill_colors) +
      ggtitle(label = paste(biological_group_, 'proportion barcodes'))

    p2 <- ggplot(df_plot,
           aes(x = standard_sample_name, y = total, fill = stock_source)) +
      ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
      geom_bar(stat = 'identity', position = position_stack(), width = 1) +
      plot_theme_() +
      scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
      scale_x_discrete(expand = c(0, 0, 0, 0)) +
      labs(fill = 'source', y = 'count') +
      scale_fill_manual(values = instock_fill_colors) +
      ggtitle(label = paste(biological_group_, 'total barcodes'))

    p3 <- ggplot(df_plot %>%
             dplyr::mutate(total = 1),
           aes(x = standard_sample_name, y = total, fill = stock_source)) +
      ggh4x::facet_nested(. ~ time_point + cell_type + organ + genetic_source, scales = 'free', space = 'free') +
      geom_bar(stat = 'identity', position = position_stack(), width = 1) +
      plot_theme_() +
      scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
      scale_x_discrete(expand = c(0, 0, 0, 0)) +
      labs(fill = 'source', y = 'count') +
      scale_fill_manual(values = instock_fill_colors) +
      ggtitle(label = paste(biological_group_, 'unique barcodes'))

    markdown_header('##', biological_group_)
    
    subchunkify(p1, fig_width = 6, fig_height = 6)

    subchunkify(p2, fig_width = 6, fig_height = 6)

    subchunkify(p3, fig_width = 6, fig_height = 6)
    
    df_stock_source_summary <- df_plot %>%
      dplyr::group_by(standard_sample_name, stock_source) %>%
      dplyr::mutate(
        unique_barcodes = dplyr::n(),
        total_barcodes = sum(total)
        ) %>%
      dplyr::ungroup() %>%
      dplyr::select(standard_sample_name, biological_group.x, cell_type, organ, time_point, stock_source, unique_barcodes, total_barcodes) %>%
      base::unique() %>%
      dplyr::rename(biological_group = biological_group.x) %>%
      tidyr::pivot_wider(names_from = 'stock_source', values_from = c('unique_barcodes', 'total_barcodes'), values_fill = 0)
    
    df_stock_source_long <- df_plot %>%
      dplyr::rename(biological_group = biological_group.x) %>%
      dplyr::select(-biological_group.y)
    
    barcode_stock_source[[biological_group_]] <- list(
      'summarized_source' = df_stock_source_summary,
      'detailed_source' = df_stock_source_long
    )

  }
  
  # writing csv files with data used to make above graphs
  
  df_source_summary <- data.frame()
  for (f in names(barcode_stock_source)) {
    df_source_summary <- dplyr::bind_rows(
      df_source_summary,
      barcode_stock_source[[f]]$summarized_source
      )
  }
  
  write.csv(df_source_summary, file.path(runs_paths$base, 'output', 'experimental_barcode_source_summary.csv'), row.names = FALSE)
  rm(df_source_summary)
  
  df_source_detailed <- data.frame()
  for (f in names(barcode_stock_source)) {
    df_source_detailed <- dplyr::bind_rows(
      df_source_detailed,
      barcode_stock_source[[f]]$detailed_source
      )
  }
  write.csv(df_source_detailed, file.path(runs_paths$base, 'output', 'experimental_barcode_source_detailed.csv'), row.names = FALSE)
  rm(df_source_detailed)
  
  } else {
  print('Only one stock source used. This section will not be processed.')
    }

```